peters_elements_causal_inference

### 2021-04-25 paper/draft.Rmd
role of time

The first section provides introduces probabilities and will focus on clarifying the difference between observations and causal concepts.
The second section discusses the hierachy of causation. 
The third section dicusses causal graphs. 
The fourth

@@

This paper focuses on atomic, policy and off-policy intervention.
Atomic intervention means that we replace a variable with a constant so we replace variable $T$ with constant $t$.
Policy intervention means we replace a variable with a new distribution $\pi$.
Off-policy intervention is an intervention when we only have historical data and no data on the specific intervention. 
Off-policy has gained traction in discplines like reinforced learning.

@@

Causal graphs are graphical models to visualize causal models. 
Causal graphs are based on various different causal models.
A graphical models contains nodes and edges (vertices).
If we are dealing with a causal graph based on a structural causal model, we usually see latent variables.
These latent variables are typically differentiated by either using a different color and/or shape than the nodes in probabilistic models where we only special conditional probabilities.
The absence of a edge between nodes can many different things depending on the underlying model.
Absence in a SCM means that there is no causal relationship.
Absence in a BCN means that there is no conditional probability.

The most common causal graph is the acyclic directed graph. 
A directed graph is a graph where every edge has a direction.
An un-directed graph has edges but there is no arrow which means that the relationship is unclear.
We can depict confounding variables via a edge with two arrows.
An acyclic graph is a graph without cycles.
That means that a variable does not cause itself, directly or indirectly.
Given the way we look at time where the relationship between variables is indifferent to time, this representation provides a foundation for graphical models.

## D-Seperation

#### In Graphical Model

#### in SCM

@@

Differential equations can describe a dynamic system accounting with time considerations.

>>>>

The first section provides introduces probabilities and will focus on clarifying the difference between observations and causal concepts.

### (Causal) Minimality
### Interventioanl equivalence 
### Cyclic SCMs
### d-seperation

A baseline SCM consists of two assingments.

\begin{equation}
A = N_A
\end{equation}
\begin{equation}
B = f_B(A, N_B)
\end{equation}

$N_A$ and $N_B$ are independent. 
A is a direct cause of B.
A-> B is causal graph (Peters et al., 2017).
The intervention distribution is unequal to the observational distribution.

**Canonical Representation**

Causal graphs are graphical models to visualize causal models. 
Causal graphs are based on various different causal models.
A graphical models contains nodes and edges (vertices).
If we are dealing with a causal graph based on a structural causal model, we usually see latent variables.
These latent variables are typically differentiated by either using a different color and/or shape than the nodes in probabilistic models where we only special conditional probabilities.
The absence of a edge between nodes can many different things depending on the underlying model.
Absence in a SCM means that there is no causal relationship.
Absence in a BCN means that there is no conditional probability.  
The most common causal graph is the acyclic directed graph. 
A directed graph is a graph where every edge has a direction.
An un-directed graph has edges but there is no arrow which means that the relationship is unclear.
We can depict confounding variables via a edge with two arrows.
An acyclic graph is a graph without cycles.
That means that a variable does not cause itself, directly or indirectly.

@@

This paper focuses on atomic, policy and off-policy intervention.
Atomic intervention means that we replace a variable with a constant so we replace variable $T$ with constant $t$.
Policy intervention means we replace a variable with a new distribution $\pi$.
Off-policy intervention is an intervention when we only have historical data and no data on the specific intervention. 
Off-policy has gained traction in discplines like reinforced learning.^[Another type of intervention is soft intervention when we only change the noise distribution and keep the prior causal effect]

@@

Given the way we look at time where the relationship between variables is indifferent to time, this representation provides a foundation for graphical models.

@@

SCMs are abstractions of physical processes.
Differential equations can describe a dynamic cyclic system accounting with more sophisticated time considerations.
Henceforth SCMs are predominately used for acyclic causal graphs.
Note that acyclic causal graphs are still dynamical models but they do not have loops where variables reinforce themselves.

### 2021-04-25 references/peters_Causality.Rmd
role of time

>>>>

**statistical vs. algorithmic independence**

# The role of Time

When interested in general cyclic system, we can use differential equations rather than SCMs as the SCM is an abstraction of underlying physical processes. 
Econometrics decided to view structural equations as algebraic equations.
Ignore interpretation as abstractions.
Natural intervention does not violate independence. 
Others may be complex.

When interested in cyclic systems differential equation unavoidable.

```{r, fig.cap="Peters et al. (2017) Ranking SCMs", include =F}
library(knitr)
include_graphics("~/Desktop/SOSE2021/causal_ml/paper/relation_physics.png")
```
